---
title: "Pitstop Data Cleaning"
author: "Hudson Double"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    number_sections: true
    highlight: tango
---
# Refined Cleaning Methods

```{r API-Setup, include=FALSE}
source("F1-API-Setup.r")
knitr::opts_chunk$set(error=FALSE)
```

## Session Isolation
```{r Session-Iscolation}
race_sessions <- dplyr::filter(session_data, session_name == "Race")
race_sessions <- race_sessions[!duplicated(race_sessions$session_key),] #Just  procoutionary, in the current state, this does nothing
```

## Driver Data Cleanup
Cleaning up the driver data into a usable state prior to use

### Merging RB and Racing Bulls
```{r Merging RB and Racing Bulls}
driver_names <- driver_names %>%
  mutate(team_name = ifelse(team_name == "RB", "Racing Bulls", team_name))
```

### Fixing Color Codes
```{r Color-Codes-Fixed}
driver_names$team_colour <- paste0("#", driver_names$team_colour)

driver_names <- driver_names %>%
  group_by(team_name) %>%
  mutate(team_colour = first(team_colour)) %>%
  ungroup()
```

### Filtering to Race Drivers Only
```{r Driver-Data}
race_drivers_all <- dplyr::filter(driver_names, session_key %in% race_sessions$session_key)
```

### Adding Driver Info
```{r Reduction-Driver-Columns}
limited_drivers <- race_drivers_all %>%
  select(-meeting_key, -broadcast_name, -first_name, -last_name)
```
## Pitstop Data Cleaning
In this section I am cleaning the pitstop_data$pit_duration to remove na values, and outliers.
### Filtering Down to Race Pitstops
```{r Race-Pitstops}
pitstop_data <- pitstop_data %>% 
  filter(session_key %in% race_sessions$session_key) 
```
### Joining Driver Information
```{r Joining}
pitstop_data <- pitstop_data %>%
  left_join(limited_drivers, by = c("session_key", "driver_number")) %>%
  filter(!is.na(pit_duration))
```

### Removing Country Codes
Removes country codes associated with dirvers from pitstop_data
```{r Removed-Country-Codes}
pitstop_data <- pitstop_data %>% 
  select(-country_code)
```

### Merging race_sessions to pitstop_data
Joining session information from race_sessions in order to track races by location and time over time. Specifically, adding Year and circuit_short_name from race_sessions to pitstop_data.
```{r Adding-Session-Info}
pitstop_data <- pitstop_data %>%
  left_join(race_sessions %>% 
  select(session_key, year, circuit_short_name, country_code), by = "session_key")
```

### Outlier Determination
- Data is tracked from the start of the pitlane to the end, regardless of condition. Under a safety car, all cars are pitted for long periods of time, same is true for a DNF.
- Due to this, I intend to remove all pitstops which exceed 150 seconds first, after which point I will remove the outliers of the remaining set.
```{r Outliers}
summary(pitstop_data$pit_duration)
pitstop_data <- pitstop_data[pitstop_data$pit_duration <= 150,]
pitstop_duration_Q1 <- quantile(pitstop_data$pit_duration, 0.25, na.rm = TRUE)
pitstop_duration_Q3 <- quantile(pitstop_data$pit_duration, 0.75, na.rm = TRUE)
pitstop_iqr <- pitstop_duration_Q3 - pitstop_duration_Q1 #I could also just do this with pitstop_iqr <- IQR(pitstop_data$pit_duration, na.rm = TRUE)
pitstop_median <- median(pitstop_data$pit_duration, na.rm = TRUE)
summary(pitstop_data$pit_duration)

# pitstop_data <- pitstop_data %>%
#   filter(pit_duration >= (pitstop_duration_Q1 - 1.5 * pitstop_iqr) & 
#          pit_duration <= (pitstop_duration_Q3 + 1.5 * pitstop_iqr))
``` 

## Pitstop Averages
```{r Pitstop Averages, include = FALSE}
pitstop_average <- pitstop_data %>%
  group_by(name_acronym, team_name) %>%
  summarise(Mean = mean(pit_duration, na.rm = TRUE))

pitstop_average <- pitstop_average %>%
  left_join(pitstop_data %>% 
              select(team_name, team_colour) %>% 
              distinct(), 
            by = "team_name")

pitstop_average <- pitstop_average %>%
  group_by(name_acronym) %>%
  mutate(driver_label = if(n() > 1) paste0(name_acronym, " (", row_number(), ")") else name_acronym) %>%
  ungroup()
```

## Plotting the Results
Using the Results I have previously gathered, this section will contain several plots created form the aggregated data points
```{r Pit-Boxplots, fig.width=12, fig.height=7, out.width='100%', out.height='600px', fig.align='center'}
TeamBoxplot <- ggplot(pitstop_data, aes(x = reorder(team_name, pit_duration, median), y = pit_duration, fill = team_colour)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.3, color = "black") +
  coord_flip() +
  scale_y_continuous(limits = c(quantile(pitstop_data$pit_duration, 0.02, na.rm = TRUE), 
                                 quantile(pitstop_data$pit_duration, 0.97, na.rm = TRUE)),
                    oob = scales::oob_keep) + #looked up this command
  labs(title = "Pit Stop Duration by Team", x = "Team", y = "Pit Stop Duration (s)") +
  theme_bw()+
  theme(legend.position = "none")+ 
  scale_fill_identity()

ggplotly(TeamBoxplot, width = 1200, height = 700) %>%
  layout(autosize = TRUE)
```


### Plotting Avergaes 
```{r Average_Plotting}
AveragesGraph <- ggplot(pitstop_average, aes(x = reorder(driver_label, Mean), y = Mean, fill = team_name))+
  geom_col()+
  scale_fill_manual(values = setNames(pitstop_average$team_colour, pitstop_average$team_name))+
  labs(x = "Name Acronyms", y = "Mean (s)", title = "Averages Plot")+
  theme(text = element_text(angle = 45, hjust = 1))

ggplotly(AveragesGraph, width = 1200, height = 700) %>%
  layout(autosize = TRUE)
```

### Mean Deviation
Scatterplot with mean_deviation trends included
```{r Mean-Deviation}
pitstop_data <- pitstop_data %>%
  group_by(session_key) %>%
  mutate(mean_deviation = pit_duration - mean(pit_duration, na.rm = TRUE)) %>%
  ungroup()
```

### Mean Deviation by Seconds
```{r Mean-Deviation-Seconds}
mean_deviation_seconds <- ggplot(pitstop_data, aes(x = pit_duration, y = mean_deviation, color = team_name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  scale_color_manual(values = setNames(pitstop_data$team_colour, pitstop_data$team_name)) +
  coord_cartesian(xlim = c(10, 45), ylim = c(-12, 18)) +
  labs(title = "Mean Deviation VS. Seconds", x = "Seconds", y = "Mean Deviation")

#ggsave("Mean_Deviation_Seconds.png", plot = mean_deviation_seconds, width = 12, height = 7, dpi = 1000)
```


### Mean Deviation by Weekend
For this section, Mean deviation is beogn tracked by weekend in order to assess team trends over time in order to asses trends in team performance. As a PNG,  this plot looks good, but it isn't readable, but in Plotly, selecting between teams will allow for more clear trends across teams. 
```{r Mean-Deviation-Weekend}
mean_deviation_weekend <- ggplot(pitstop_data, aes(x = interaction(circuit_short_name, year, sep = " "), y = mean_deviation, color = team_name)) +
  annotate("Rect", xmin = -Inf, xmax = Inf, ymin = 0, ymax = Inf, fill = "red", alpha = 0.1)+
  geom_point(size = 0.3) +
  geom_smooth(aes(group = team_name), method = "lm", se = FALSE) +
  coord_cartesian(ylim = c(-2,2)) +
  scale_color_manual(name = "Team Name", values = setNames(pitstop_data$team_colour, pitstop_data$team_name)) +
  labs(title = "Mean Deviation By Weekend", x = "Circuit and Year", y = "Mean Deviation") +
  theme_minimal() +
  scale_y_continuous(breaks = seq(-5, 5, by = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplotly(mean_deviation_weekend)

#ggsave("Mean_Deviation_Weekend.png", plot = mean_deviation_weekend, width = 12, height = 7, dpi = 1000)
```